Forecasting the 2024 U.S. Presidential Election

### 2024-11-02 paper/paper.qmd
Add more details in Model section

Below, we briefly describe the multiple linear regression model used to examine Trump’s winning probability. Additional details and model diagnostics are provided in [Appendix -@sec-model-details].

## Multiple Linear Regression Model Overview
The model now predicts Trump’s polling percentage (percent) using the following predictors:

@@

```{r}
#| echo: false
#| label: tbl-summary-model
#| tbl-cap: Regression Model Summary Table (Display the first 10 rows)
#| warning: false
#| message: false

regression_model<-readRDS(here::here("models/regression_model.rds"))
summary<-summary(regression_model)
# Extract the coefficients to form a table
coefficients_table <- summary$coefficients[, 1:3]

# Display the first 10 rows of the coefficients table using kable, formatted to 2 decimal places
kable(head(coefficients_table, 10), digits = 2)
```

@@

# Create the table using kable
kable(state_predictions[, c("state", "predicted_percent_trump", "electoral_votes", "winner")], 
      col.names = c("State", "Trump Predicted %", "Electoral Votes", "Winner"), 
      digits = 2)
```

@@

## Weakness
Despite the model estimating the voter's choice trend, it has limitations. For instance, assuming linear relationships between predictors may oversimplify complex voter behaviors and other external social influences. The dataset simplifies the potential influence behind the US election.

@@

# Display the full summary table using kable
kable(full_summary_df)

>>>>

One alternative model considered is logistic regression, commonly used for binary outcomes to predict the probability of an event occurring versus not occurring. While logistic regression is effective for classification tasks, it is not suitable here, as the goal of this paper is to predict a continuous percentage of voters likely to vote for Donald Trump. Logistic regression restricts outputs to probabilities between 0 and 1, which does not align with the requirements of this analysis. Therefore, a multiple linear regression framework is selected for this paper.

## Multiple Linear Regression Model Overview
Key assumptions for Multiple linear regression:

- Linearity: The relationship between the independent variables (predictors) and the dependent variable(Percentage of Respondents Supporting Donald Trump) is linear. 
- Independence: The outcome of one observation(respondents) should not influence another
- Homoscedasticity: The variance of the residuals (errors) is constant across all levels of the independent variables.
- Normality of Residuals: The residuals (errors) of the model are normally distributed.
- No Multicollinearity: Independent variables should not be highly correlated with each other. Multicollinearity will make some predictors appear insignificant when they might actually have an effect.

Additional details and model diagnostics are provided in [Appendix -@sec-model-details].

The model predicts Trump’s polling percentage (percent) using the following predictors:

@@

```{r}
#| echo: false
#| label: tbl-summary-model
#| tbl-cap: Regression Model Summary Table. This table displays the top 10 predictors with the largest absolute coefficients, along with Transparency Score and Days Until Election. The standard errors represent the variability in the coefficient estimates, while the t-values indicate the strength and significance of each predictor’s relationship with the outcome variable.
#| warning: false
#| message: false

# Load the regression model
regression_model <- readRDS(here::here("models/regression_model.rds"))
summary <- summary(regression_model)

# Extract the coefficients and form a table with estimates, standard errors, and t-values
coefficients_table <- as.data.frame(summary$coefficients[, 1:3])

# Select the top 10 rows with the largest absolute coefficients
coefficients_table$Predictor <- rownames(coefficients_table)
coefficients_table <- coefficients_table[order(abs(coefficients_table$Estimate), decreasing = TRUE), ]
top_10_table <- head(coefficients_table, 10)

# Ensure that transparency_score and days_until_election are included in the table
required_predictors <- c("transparency_score", "days_until_election")
additional_predictors <- coefficients_table[coefficients_table$Predictor %in% required_predictors, ]

# Combine the top 10 table with the required predictors, ensuring no duplicates
final_table <- unique(rbind(top_10_table, additional_predictors))

# Rename specific predictors
final_table$Predictor <- as.character(final_table$Predictor)
final_table$Predictor[final_table$Predictor == "transparency_score"] <- "Transparency Score"
final_table$Predictor[final_table$Predictor == "days_until_election"] <- "Days Until Election"

# Set row names to the updated Predictor names and remove the Predictor column
row.names(final_table) <- final_table$Predictor
final_table <- final_table[, -ncol(final_table)]  # Remove Predictor column

# Rename the columns
colnames(final_table) <- c("Coefficient", "Standard Error", "t-Statistic")

# Display the table
kable(final_table, digits = 2)

@@

# Create the table using kable
kable(state_predictions[, c("state", "predicted_percent_trump", "electoral_votes", "winner")],        col.names = c("State", "Trump Predicted %", "Electoral Votes", "Winner"),digits = 2) 
```

@@

## Weakness
Despite the model’s ability to estimate voter choice trends, it has several limitations. For instance, assuming linear relationships between predictors restricts its capacity to capture more complex, non-linear relationships, potentially oversimplifying external social influences. The dataset itself may also oversimplify the range of factors affecting the U.S. election, limiting the model’s ability to account for all relevant influences. Additionally, outliers and influential points can disproportionately impact the regression model, potentially distorting results and leading to misleading conclusions.

@@

```
```{r}
#| echo: false
#| tbl-cap: Coefficient Estimates with 95% Confidence Intervals for Predictors. Predictors with positive coefficients are shown in blue, while predictors with negative coefficients are shown in red. 
#| warning: false
#| message: false

# Convert the model summary to a tidy format for easier plotting
tidy_model <- broom::tidy(regression_model, conf.int = TRUE, conf.level = 0.95) %>%
  filter(term != "(Intercept)")

# Plot the coefficients with confidence intervals, excluding the intercept
ggplot(tidy_model, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(aes(color = estimate > 0), size = 1.2) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), color = "grey", height = 0.3) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  scale_color_manual(
    values = c("red", "blue"),
    labels = c("Negative Coefficient", "Positive Coefficient"),
    name = ""  # This removes the default "estimate > 0" label
  ) +
  labs(x = "Coefficient Value", y = "Predictors") +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y = element_text(size = 3),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    plot.title = element_text(size = 16, face = "bold"),
    legend.position = "top"
  ) +
  guides(color = guide_legend(reverse = TRUE))
